# Import libraries
```{r}
library(sf)
library(sp)
library("tidyverse")
library("purrr")
library("tidycensus")
library("tigris")
library(readxl)
library(here)
library(codebook)
library(labelled)
library('fastDummies')
```


# Read an existing API key
```{r}
API_KEY <- Sys.getenv("CENSUS_API_KEY")
census_api_key(API_KEY)
```


# Load zcta, census block, and census tract dataframes
```{r}
census_api_key(API_KEY)

years <- lst(2019)
v19 <- load_variables(2019, "acs5", cache = TRUE)

tract2019_frame <- map_dfr(
  years,
  ~ get_acs(
    geography = "tract",
    variables = c(pop="B01001_001" , #Total population
                  housing="B11001_001", #Total number of housing units
                  white = "B11001A_001" #Total number of householders who are White alone
    ),
    state = c("DE"),
    year = .x,
    geometry = TRUE,
    output="tidy"
  ),
  .id = "year"  # when combining results, add id var (name of list item)
)

years <- lst(2020)
v20 <- load_variables(2020, "acs5", cache = TRUE)

tract2020_frame <- map_dfr(
  years,
  ~ get_acs(
    geography = "tract",
    variables = c(pop="B01001_001" , #Total population
                  housing="B11001_001", #Total number of housing units
                  white = "B11001A_001" #Total number of householders who are White alone
    ),
    state = c("DE"),
    year = .x,
    geometry = TRUE,
    output="tidy"
  ),
  .id = "year"  # when combining results, add id var (name of list item)
)

years <- lst(2020)
v20_dec <- load_variables(2020, "pl", cache = TRUE)

block_frame <- map_dfr(
  years,
  ~ get_decennial(
    geography = "block",
    variables = c(pop="P1_001N" # Use population as weighting variable
                  #pop="H1_001N" # Use number of houses as weighting variable
                  ),
    state = c("DE"),
    year = .x,
    sumfile = "pl",
    geometry = TRUE,
    output="tidy"
  ),
  .id = "year"  # when combining results, add id var (name of list item)
)
```

# Rename column from 'value'
```{r}
block_frame <- block_frame %>% rename(the_value=value)
```

# Collapse dataframes to one entry per GEOID
```{r}
# Collapse tract2019 data
tract2019_pop_frame <- tract2019_frame %>% filter(variable == "pop") %>% rename(pop=estimate) %>% select(GEOID, pop, geometry)
tract2019_house_frame <- tract2019_frame %>% filter(variable == "housing") %>% rename(housing=estimate) %>% select(GEOID, housing) %>% st_drop_geometry()
tract2019_white_frame <- tract2019_frame %>% filter(variable == "white") %>% rename(white=estimate) %>% select(GEOID, white) %>% st_drop_geometry()
merged_tract2019_frame <- merge(tract2019_pop_frame, tract2019_house_frame, by="GEOID") %>% merge(tract2019_white_frame, by="GEOID")

# Collapse tract data
tract2020_pop_frame <- tract2020_frame %>% filter(variable == "pop") %>% rename(pop=estimate) %>% select(GEOID, pop, geometry)
tract2020_house_frame <- tract2020_frame %>% filter(variable == "housing") %>% rename(housing=estimate) %>% select(GEOID, housing) %>% st_drop_geometry()
tract2020_white_frame <- tract2020_frame %>% filter(variable == "white") %>% rename(white=estimate) %>% select(GEOID, white) %>% st_drop_geometry()
merged_tract2020_frame <- merge(tract2020_pop_frame, tract2020_house_frame, by="GEOID") %>% merge(tract2020_white_frame, by="GEOID")
```

# Get dummy variables for zipcode of origin
```{r}
merged_tract2019_frame <- dummy_cols(merged_tract2019_frame, select_columns = "GEOID")
merged_tract2019_frame <- st_as_sf(merged_tract2019_frame)
```


# Interpolate data onto census tracts using population weights from census blocks
```{r}
interpolated <- interpolate_pw(
  merged_tract2019_frame,
  merged_tract2020_frame,
  to_id = "GEOID",
  extensive = TRUE, 
  weights = block_frame,
  weight_column = "the_value",
  crs = 26949
)
```


# Write the interpolated data to a csv
```{r}
## Save the summary list
write_csv(interpolated,
          here("interpolated.csv") #Population weighted interpolation
          #here("interpolated_house.csv") #House weighted interpolation
          )
```

# Write the census tract data from 2020 to a csv
```{r}
## Save the summary list
write_csv(merged_tract2020_frame,
          here("tracts2020.csv")
          )
```


# Acquire an API key
```{r}
API_KEY <- Sys.getenv("API_KEY")
census_api_key(API_KEY,install=TRUE,overwrite=TRUE)
```
